literature coding explore

### 2024-04-20 reports/lit_codes_explore.qmd
add w/in code frequencies plots

### Code appearance frequencies within code categories

>>>>

### Code appearance frequencies within code categories
Here, I have plotted the same information as the previous section, but as a percentage of the overall bridges and barriers groupings. For example, data bridges make up 37% of all bridge codes. The same pattern holds for the other sub-panels. For example, on the ecological page, abiotic indicators account for 22% of all ecological bridge codes.

::: panel-tabset

#### Broader categories
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 10
#| fig-height: 7
#| fig-align: center

code_sum <- code_freq %>%
  group_by(bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(code_cat) %>% 
  mutate(n_subtotal = sum(total)) %>%
  summarise(value = n_subtotal/n_total) %>%
  rename("br_ba" = code_cat) %>%
  mutate(percent_label = scales::percent(value, accuracy = 1), 
         percent = if_else(br_ba == "Data barrier" | br_ba == "Ecological barrier" | br_ba == "Social barrier", value, -value),
         code_level = if_else(br_ba %in% c("Ecological barrier", "Ecological bridge"), "Ecological bridge or barrier", 
                              if_else(br_ba %in% c("Data barrier", "Data bridge"), "Data bridge or barrier",
                              if_else(br_ba %in% c("Social barrier", "Social bridge"), "Social bridge or barrier", NA))), 
         code_brba = if_else(br_ba %in% c("Data bridge", "Ecological bridge", "Social bridge"), "bridge", "barrier")) %>%
  group_by(code_level) %>%
  distinct()

code_sum$code_level <- factor(code_sum$code_level, levels = c("Social bridge or barrier", "Ecological bridge or barrier", "Data bridge or barrier"))
code_sum$code_brba <- factor(code_sum$code_brba, levels = c("bridge", "barrier"))

ggplot(code_sum, aes(x = code_level, 
             y = percent,
             fill = code_brba)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = percent_label),
            position = position_stack(vjust = 0.65),
            color = "black", 
            size = 5) +
  ylim(-0.4,0.7)+
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"), labels = c("Bridge frequency", "Barrier frequency")) +
  theme_minimal() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x =element_blank(),
        axis.text.y = element_text(size = 20, color = "black"),
        panel.grid = element_blank(),
        legend.position = c(-0.17, 0.99),
        legend.text = element_text(size = 16, color = "black"),
        strip.text.y.right = element_text(angle = 0, size = 16)) +
  geom_hline(yintercept = 0.0, linewidth = 1) + 
  labs(fill = "")+ 
  guides(fill = guide_legend(nrow = 1))

#### Data codes
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

data_freq <- code_freq %>%
  filter(code_cat == "Data bridge" | code_cat == "Data barrier")

data_sum <- data_freq %>%
  group_by(bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(code_name) %>% 
  summarise(value = total/n_total) %>%
  mutate(code_cat = ifelse(code_name == "Implementation costs" | code_name == "Gaps and inaccuracy" | code_name == "Management scales do not align with system needs" | code_name == "Model abuse" | code_name == "Climate change adding uncertainty to target species habitat use", "Data barrier", "Data bridge")) %>%
  mutate(percent_label = scales::percent(value, accuracy = 2))

my_labs <- c("0", "25", "50", "75")
data_sum$code_cat <- factor(as.factor(data_sum$code_cat), levels = c("Data bridge", "Data barrier"))

ggplot(data_sum) +
  geom_bar(aes(x = value, y = reorder(code_name, value), fill = code_cat), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,0.9),
                     expand = c(0,0), 
                     breaks = c(0, 0.25, 0.5, 0.75),
                     labels = my_labs)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = value, y = code_name, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent from all data bridges or barriers (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~code_cat, scales = "free", nrow = 2)

#### Ecological codes
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

eco_freq <- code_freq %>%
  filter(code_cat == "Ecological bridge" | code_cat == "Ecological barrier")

eco_sum <- eco_freq %>%
  group_by(bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(code_name) %>% 
  summarise(value = total/n_total) %>%
  mutate(code_cat = ifelse(code_name == "Need for multispecies management", "Ecological barrier", "Ecological bridge")) %>%
  mutate(percent_label = scales::percent(value, accuracy = 2))

eco_sum$code_cat <- factor(as.factor(eco_sum$code_cat), levels = c("Ecological bridge", "Ecological barrier"))
my_labs2 <- c("0", "30")

ggplot(eco_sum) +
  geom_bar(aes(x = value, y = reorder(code_name, value), fill = code_cat), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,0.35),
                     expand = c(0,0), 
                     breaks = c(0, 0.3),
                     labels = my_labs2)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = value, y = code_name, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent from all ecological bridges or barriers (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~code_cat, scales = "free", nrow = 2)

#### Social codes
```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| fig-align: center

soc_freq <- code_freq %>%
  filter(code_cat == "Social bridge" | code_cat == "Social barrier")

soc_sum <- soc_freq %>%
  group_by(bridge_barr) %>%
  mutate(n_total = sum(total)) %>%
  ungroup() %>%
  group_by(code_name) %>% 
  summarise(value = total/n_total) %>%
  mutate(code_cat = ifelse(code_name == "Indirect socioeconomic consequences" | code_name == "Lack of political administration permanence" | code_name == "Poor communication and outreach" | code_name == "Poor program implementation", "Social barrier", "Social bridge")) %>%
  mutate(percent_label = scales::percent(value, accuracy = 2))

soc_sum$code_cat <- factor(as.factor(soc_sum$code_cat), levels = c("Social bridge", "Social barrier"))
my_labs2 <- c("0", "30")

ggplot(soc_sum) +
  geom_bar(aes(x = value, y = reorder(code_name, value), fill = code_cat), width = 0.6, stat = "identity")+
  scale_x_continuous(position = "top", 
                     limits = c(0,1),
                     expand = c(0,0), 
                     breaks = c(0,0.25, 0.5, 0.75),
                     labels = my_labs)+
  scale_y_discrete(expand = expansion(add = c(0,0))) +
  tidyquant::theme_tq() +
  theme(
    # Set background color to white
    panel.background = element_rect(fill = "white"),
    # Set the color and the width of the grid lines for the horizontal axis
    panel.grid.major.x = element_line(color = "#A8BAC4", size = 0.3),
    # Remove tick marks by setting their length to 0
    axis.ticks.length = unit(0, "mm"),
    # Remove the title for both axes
    axis.title = element_blank(),
    # Only left line of the vertical axis is painted in black
    axis.line.y.left = element_line(color = "black"),
    # But customize labels for the horizontal axis
    axis.text.x = element_text(family = "Econ Sans Cnd", size = 16), 
    axis.text.y = element_text(family = "Econ Sans Cnd", size = 16), 
    legend.position = "none", 
    strip.text = element_text(size = 16)) + 
  geom_text(aes(x = value, y = code_name, label = percent_label),
    hjust = 0,
    nudge_x = -0.0001,
    colour = "black",
    family = "Econ Sans Cnd",
    size = 6)+
  labs(subtitle = "Percent from all social bridges or barriers (%)") + 
  theme(plot.subtitle = element_text(
      family = "Econ Sans Cnd",
      size = 20)
  ) + 
  scale_fill_manual(values = c("goldenrod", "lightskyblue3"))+
  facet_wrap(~code_cat, scales = "free", nrow = 2)
